---
title: "Basic Inferential Data Analysis"
author: "Marçal Boix"
date: "Nov 21, 2015"
output: pdf_document
latex_engine: xelatex
---

In the second part of the project, we analyze the ToothGrowth data in the R datasets package. The data is set of 60 observations, length of odontoblasts (teeth) in each of 10 guinea pigs at each of three dose levels of Vitamin C (0.5, 1 and 2 mg) with each of two delivery methods (orange juice or ascorbic acid).

## Initialization
In order to complete the project we need to initialize R with a few statements.
```{r, echo=TRUE}
library(ggplot2)
library(pander)
```

## Loading data
```{r, echo=TRUE}
# ToothGrowth - The Effect of Vitamin C on Tooth Growth in Guinea Pigs.
library(datasets)
head(ToothGrowth)
```

## Basic Summary
```{r, echo=TRUE}
# Structure of data frame.
str(ToothGrowth)
# Summary of columns. Few statistics for each column.
summary(ToothGrowth)
# Number of observations by supplement type and dose.
table(ToothGrowth$supp, ToothGrowth$dose)
```

## Correlation
```{r}
ggplot(data=ToothGrowth, aes(x=as.factor(dose), y=len, fill=supp)) +
    geom_bar(stat="identity") +
    facet_grid(. ~ supp) +
    xlab("Dose in miligrams") +
    ylab("Tooth length") +
    guides(fill=guide_legend(title="Supplement type")) +
    theme_bw()
```

As we can see in the graph above, there is a positive correlation between the tooth length and the dose levels of Vitamin C, for both delivery methods.  

For each supplement type and dose, we will calculate mean, sd and variance.
```{r}
dose = as.numeric(levels(as.factor(ToothGrowth$dose)))
supp = levels(ToothGrowth$supp)
basicSt = list()
for (d in dose)
{
    for (s in supp)
    {
        subTG = ToothGrowth$len[ToothGrowth$dose==d & ToothGrowth$supp==s]
        basicSt <<- rbind(basicSt, list(supp = s, dose = d, mean=mean(subTG), 
                                        sd=sd(subTG), var =var(subTG)))
    }
}
as.data.frame(basicSt)
```

## Multiple hypothesis testing
To check if there is a real difference between the groups by dose level and delivery method, we will do a two-sided unpaired t-tests.  
This test allow us to obtain the confidence intervals and p-values.  
The null hypothesis, in all cases, is that there is no difference in the means between the two groups.
```{r}
tests = list()
for (d in dose) {
  ojDose <- ToothGrowth$len[ToothGrowth$dose == d & ToothGrowth$supp == "OJ"]
  vcDose <- ToothGrowth$len[ToothGrowth$dose == d & ToothGrowth$supp == "VC"]
  # We assume unequal variance.
  t <- t.test(ojDose, vcDose, var.equal=FALSE, paired=FALSE)
  id <- paste("OJ", d, "-", "VC", d)
  tests <- rbind(tests, list(id=id, p.value=round(t$p.value,5), 
                             ci.lo=round(t$conf.int[1],6), ci.hi=round(t$conf.int[2],6)))
}
tests
```

## Conclusions
#. For 0.5 and 1 mg dose levels, there is a significant difference between the means of the OJ and VC groups (p-values < 0.05  and the 95% confidence intervals doesn't include zero).
#. For the 2 mg dose level, we fail to reject the null hypothesis  (p-value > 0.5 and the 95% confidence interval includes zero). There is no significative influence of the delivery method on tooth growth in guinea pigs, for 2mg dose.
